---
title: "Second assignment"
author: "Asma"
date: '2022-06-20'
output: html_document
---
## 1- Introduction about our data set 
## 2- Describe the attributes in the dataset 
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/misk-DSI")
Data <- read.csv("coffee_ratings.csv")
e <- sum(is.na(Data)) 
#na_av <- (e/nrow(Data))*100
## delete all NA value from the dataset =(132!)rows 
#Data2 <-na.omit(Data)
nrow(Data)
str(Data)
summary(Data)
# this function calculates the percentage of NA in each column
p <- function(x){round(sum(is.na(x)/length(x)*100),digits = 2)}
#apply the previous function to the data 
apply(Data ,2,p)
#--------- can't remember exactly this code for what ----------
#library(mice)
#md.pattern(Data,c(""))
#md.pairs(Data)
#---------------------------------------------------------

```


## 3- Describe the procedures we follow to clean the data 

### 1- Delete the columns and rows which contain NA value


```{r}
 
# Delete the column if it contain more than (500) NA value 
# for (r in 1:ncol(Data)){
#   if (sum(is.na(Data[,r]))>500){
#     Data=Data[,-r]}
# }

# Delete the column if it contain more than (?) NA value

#for (r in 1:nrow(Data)){
 # if (sum(is.na(Data[r,]))>10){
  #  Data3=Data[-r,]}
#}


```

### 2- Replace the NA - the numeric value by mean and the characters by mode


```{r}
#create the function to calculate the mean and mode depend on data type 
getimputation = function(x) {
  if (is.numeric(Data[,i])==TRUE){
    mean(x, na.rm = TRUE)}
  else{
  uniq = unique(x)
  uniq[which.max(tabulate(match(x, uniq)))]
  }}

# apply the function on our data 

for(i in 1:(ncol(Data))){
    Data[is.na(Data[,i]), i] = getimputation(Data[,i])
}
```



## create the world map-1

```{r}
library("ggplot2")
WorldMap = map_data('world')
#unique(WorldMap$region)
M = ggplot()
M = M + geom_map(data=WorldMap, map=WorldMap,
                  aes(x=long, y=lat, group=group, map_id=region),
                  fill="white", colour="#7f7f7f", size=0.5)
M = M + geom_map(data=Data, map=WorldMap,
                  aes(fill= total_cup_points, map_id= country_of_origin , color=total_cup_points , alpha = 0.7))
                 #,
                  #colour="#7f7f7f", size=0.5)
#M = M + scale_fill_continuous(low="thistle1", high="darkred",
 #                              guide="colorbar")
M = M + scale_y_continuous(breaks=c())
M = M + scale_x_continuous(breaks=c())
Legend=c("-----------")
M = M + labs(fill=Legend, title="", x="", y="")
M = M + theme_bw()
M = M + theme(panel.border = element_blank(),
               panel.background = element_rect(fill = '#DBFFFF'),
              legend.background = element_blank(),
               legend.position = c(0.05, 0.2))
M
```



##  Study the relashinship between   processing_method and acidity 

```{r}






ggplot(data=subset(Data , processing_method %in%c("Washed / Wet","Natural / Dry" ,"Semi-washed / Semi-pulped" ,"Pulped natural / honey")) ,aes(processing_method,acidity ))+
  geom_col(fill="lightblue"  )



```

##  Study the relashinship between   color and acidity 


```{r}

library(ggplot2) 
ggplot(data=subset(Data,color %in% c("Bluish-Green" ,"Blue-Green" ,"Green")),aes(color,acidity))+
  geom_col( fill="lightblue")
# 
```


## Study the Correlation  between  moisture and aroma


```{r}
# library(ggplot2) 
# ggplot(Data ,aes(moisture,aroma))+
#   geom_col()

plot(moisture ~ aroma ,data = Data , main="Study the relashinship between  moisture and aroma" )
cor(Data$moisture , Data$aroma , use="complete.obs") #-0.1341272
```


## Study the Correlation  between  moisture and acidity

```{r}
# library(ggplot2) 
# ggplot(Data ,aes(moisture,acidity))+
#   geom_col()

plot(moisture ~ acidity ,data = Data , main="Study the relashinship between  moisture and acidity"  )
cor(Data$moisture , Data$acidity , use="complete.obs") #-0.1290795
```



## Study the Correlation  between  acidity and aroma 

```{r}
# library(ggplot2) 
# ggplot(Data ,aes(acidity , aroma))+
#   geom_col()

plot(acidity ~ aroma ,data = Data , main="Study the relashinship between acidity and aroma"  )
cor(Data$acidity , Data$aroma , use="complete.obs") #0.7235822
```




## Study the Correlation  between  acidity and flavor 


```{r}
plot(acidity ~ aroma ,data = Data , main="Study the relashinship between acidity and flavor "  )
cor(Data$acidity , Data$flavor , use="complete.obs") #0.8166883
```











